### 2023-06-13 tex/sec4-bidirectionalisation.tex
[ tex ] Split BiSig.tex into smaller files

%! TEX root = BiSig.tex

\section{Soundness, Completeness, and Bidirectionalisation} \label{sec:annotatability}

\subsection{Soundness and Completeness}
\begin{theorem}[Soundness]\label{thm:term-soundness}

\end{theorem}

\begin{theorem}[Completeness]\label{thm:term-completeness}

\end{theorem}

\subsection{Bidirectionalisation}
\begin{proposition}[Decidability for Bidirectionalisation] \label{thm:bidirectionalisation}
  For every raw term $t$ and a mode $d$, it is decidable that $t$ is in mode $d$, i.e.\ $\vdash_{\Sigma, \Omega} \isTerm{t}^\isDir{d}$ is derivable.
\end{proposition}

\subsection{Completeness versus Annotatability}