### 2023-06-14 tex/sec4-bidirectionalisation.tex
[ tex ] Fix the colour of modes; use the verb form of decidable.
For example, instead of saying “for every type A it is decidable whether …”, we can say “every type A decides whether …”

%!TEX root = BiSig.tex

\section{Soundness, Completeness, and Bidirectionalisation} \label{sec:annotatability}

\subsection{Soundness and Completeness}
\begin{theorem}[Soundness]\label{thm:term-soundness}

\end{theorem}

\begin{theorem}[Completeness]\label{thm:term-completeness}

\end{theorem}

\subsection{Bidirectionalisation}
\begin{proposition}[Decidability for Bidirectionalisation] \label{thm:bidirectionalisation}
  It is decidable whether a raw term~$t$ is in mode~$d$ for any $t$ and $d$.
\end{proposition}

\subsection{Annotatability is not Completeness}