### 2023-06-16 tex/sec7-future.tex
[ tex ] Writing job distribution

%!TEX root = BiSig.tex

\section{Concluding Remarks} \label{sec:future}

\Josh{Getting rid of the assumption of being syntax-directed (the triangular picture) --- 0.5p}
\Josh{Ornaments -- 0.25p}
\Josh{NDGP -- 0.25p (optional?), \citep{Ko2022} }

\LT{polymorphic algebraic theories, dependent signatures --- 0.5p}

\cite{Fiore2013,Hamana2011,Hamana2022}