### 2023-06-25 tex/sec4-pre-synthesis.tex
[ tex ] More proofs in S4

%!TEX root = BiSig.tex

\section{Soundness, Completeness, and Mode Preprocessing}\label{sec:pre-synthesis}

\Josh{Overview of the section}

\subsection{Soundness and Completeness}
\label{sec:soundness-and-completeness}

%\Josh{Throughout this sub-section, we quantify over any bidirectional type system~$(\Sigma, \Omega)$, context $\Gamma : \Cxt_\Sigma(\emptyset)$, raw term $\erase\Gamma \vdash_{\Sigma, \erase\Omega} t$, mode~$d$, and type $A : \Type_\Sigma(\emptyset)$.}

Erasure of a bidirectional binding signature removes mode information and keeps everything else intact; this can be straightforwardly extended by induction to remove mode information from a bidirectional typing derivation and arrive at an ordinary typing derivation.

\begin{lemma}[Soundness]\label{thm:term-soundness}
If\/ $\Gamma \vdash_{\Sigma, \Omega} t :^\dir{d} A$, then $\Gamma \vdash_{\Sigma, \erase\Omega} t : A$.
\end{lemma}

\begin{proof}
Induction on the given derivation, mapping every bidirectional typing rule to its mode-less counterpart except $\ChkRule{Sub}$, in which case the induction hypothesis $\Gamma |-_{\Sigma,\erase\Omega} t : B$ suffices due to the premise $B = A$.
\end{proof}

We can also remove typing and retain mode information, arriving at a mode derivation instead.

\begin{proposition}\label{thm:typing-removal}
If\/ $\Gamma \vdash_{\Sigma, \Omega} t :^\dir{d} A$, then $\erase\Gamma \vdash_{\Sigma, \Omega} t^\dir{d}$.
\end{proposition}

\begin{proof}
Induction on the given derivation, mapping every rule to its counterpart.
\end{proof}

Conversely, if we have both mode and typing derivations for the same term, we can combine them and obtain a bidirectional typing derivation.

\begin{lemma}[Completeness]\label{thm:term-completeness}
If\/ $\erase\Gamma |-_{\Sigma, \Omega} t^\dir{d}$ and\/ $\Gamma |-_{\Sigma, \erase\Omega} t : A$, then $\Gamma |-_{\Sigma, \Omega} t :^\dir{d} A$.
\end{lemma}

\begin{proof}
Induction on the given mode derivation.
For \SynRule{Var}, \SynRule{Anno}, and \Rule{Op}, the outermost rule used in the given typing derivation must be the corresponding typing rule, so by the induction hypotheses we have bidirectional typing derivations for all the sub-terms, to which we can then apply the corresponding bidirectional typing rule.
The \ChkRule{Sub} case is similar but slightly simpler: the induction hypothesis directly gives us a derivation of $\Gamma |-_{\Sigma,\Omega} t \syn A$, to which we apply $\ChkRule{Sub}$.
\end{proof}

In short, soundness and completeness are no more than the separation and combination of mode and typing information carried by the three kinds of derivations while keeping their basic structure, which is directed by the same raw term.

\subsection{Mode Preprocessing}
\label{sec:mode-preprocessing}

\newcommand{\True}{\mathbf{T}}
\newcommand{\False}{\mathbf{F}}

\begin{figure}
  \centering
  \small
  \begin{tabular}{ r r l }
    & & is in (generalised) mode~$d$, \\
    $\smash{\boxed{V |-_{\Sigma, \Omega} \isTerm{t}^{\dir{d}\,v\,e}}}$
    & The raw term~$\isTerm{t}$\hspace{-.6em}
    & misses some type annotation iff $v = \False$, and \\
    & & is in (generalised) mode~$d$ due to an outermost mode cast iff $e = \False$
  \end{tabular}
  \begin{mathpar}
    \inferrule{x \in V}{V |-_{\Sigma, \Omega} \isTerm{x}^{\syn\,\True\,\True}}\,\SynRule{Var}
    \and
    \inferrule{\cdot |-_{\Sigma} A \\ V |-_{\Sigma, \Omega}\isTerm{t}^{\chk\,v\,e}}{V |-_{\Sigma, \Omega} (\isTerm{t \annotate A})^{\syn\,v\,\True}}\,\SynRule{Anno}
    \and
    \inferrule{V |-_{\Sigma, \Omega}\isTerm{t}^{\chk\,v\,\True}}{V |-_{\Sigma, \Omega} \isTerm{t}^{\syn\,\False\,\False}}\,\SynRule{Missing}
    \and
    \inferrule{V |-_{\Sigma, \Omega} \isTerm{t}^{\syn\,v\,\True}}{V |-_{\Sigma, \Omega} \isTerm{t}^{\chk\,v\,\False}}\,\ChkRule{Sub}
    \and
    \inferrule{V, \vec x_1 |-_{\Sigma, \Omega} \isTerm{t_1}^{\dir{d_1}\,v_1\,e_1} \\ \cdots \\ V, \vec x_n |-_{\Sigma, \Omega} \isTerm{t_n}^{\dir{d_n}\,v_n\,e_n}}
    {V |-_{\Sigma, \Omega} \tmOpts^{\dir{d}\,(\bigwedge_i v_i)\,\True}}\,\Rule{Op}
  \end{mathpar}
  \caption{Generalised mode derivations}
\end{figure}

\begin{lemma}\label{thm:adjustment}
For any raw term $V |-_{\Sigma,\erase\Omega} t$, if
\begin{equation}\label{eq:classification}
V |- t^{\syn\,\True\,\True} ~~\vee~~ V |- t^{\chk\,\True\,\True} ~~\vee~~ \exists e.~V |- t^{\chk\,\False\,e}
\end{equation}
then for any mode~$\dir{d}$, there is a derivation of\/ $V |-_{\Sigma,\Omega} t^{\dir{d}\,v\,e}$ for some booleans $v$~and~$e$.
\end{lemma}

\begin{proof}
Case analysis on statement~(\ref{eq:classification}) and~$\dir{d}$.
%\begin{itemize}
%\item For the first disjunct with $\dir{d} = {\syn}$, the second disjunct with $\dir{d} = {\chk}$, and the third disjunct with $\dir{d} = {\chk}$:
%The mode we have matches the mode to be established, so no adjustment is needed.
%\item For the first disjunct with $\dir{d} = {\chk}$:
%Apply $\ChkRule{Sub}$.
%\item For the second disjunct with $\dir{d} = {\syn}$:
%Apply $\SynRule{Missing}$.
%\item For the third disjunct with $\dir{d} = {\syn}$:
%If $e = \False$, then the outermost rule of the derivation must be $\ChkRule{Sub}$, removing which we get a derivation of $V |-_{\Sigma,\Omega} t^{\syn\,\False\,\True}$;
%otherwise $e = \True$, in which case we can apply $\SynRule{Missing}$.
%\vspace{-\topsep-\baselineskip}
%\end{itemize}
\end{proof}

Intuitively, statement~(\ref{eq:classification}) gives a precise classification of a term as either a `real' synthesising term (i.e.~without missing annotations or an outermost mode cast), a real checking term, or a term that has missing annotations.
The lemma performs some slight adjustment that involves only outermost mode casts to switch an input derivation to any required mode.

\begin{theorem}[Generalised Mode Preprocessing]\label{thm:generalised-mode-preprocessing}
For any raw term $V |-_{\Sigma,\erase\Omega} t$ and mode~$\dir{d}$, there is a derivation of\/ $V |-_{\Sigma,\Omega} t^{\dir{d}\,v\,e}$ for some booleans $v$~and~$e$.
\end{theorem}

\begin{proof}
It suffices to prove statement~(\ref{eq:classification}) by induction on~$t$ and then apply \cref{thm:adjustment}.
\begin{itemize}
\item $\Rule{Var}$:
A variable is a real synthesising term, so we establish the first disjunct using $\SynRule{Var}$.
\item $\Rule{Anno}$:
Let $t = (t' \bbcolon A)$.
By the induction hypothesis and \cref{thm:adjustment}, there is a derivation of $V |-_{\Sigma,\Omega} {t'}^{\chk\,v\,e}$ for some booleans $v$~and~$e$.
If $v = \False$, then $t'$, and therefore~$t$, have missing annotations, and we establish the third disjunct using $\SynRule{Anno}$ and $\ChkRule{Sub}$; otherwise, $t$~is a real synthesising term, and we establish the first disjunct using $\ChkRule{Sub}$.
\item $\Rule{Op}$:
Let $t = \tmOpts$.
By the induction hypotheses and \cref{thm:adjustment}, there is a derivation of $V, \vec x_i |-_{\Sigma,\Omega} {t_i}^{\dir{d_i}\,v_i\,e_i}$ for each sub-term~$t_i$; apply $\Rule{Op}$ and we get a derivation of $V |-_{\Sigma,\Omega} t^{\dir d\,v\,\True}$ where $v = \bigwedge_i v_i$.
If $v = \True$, then $t$~is a real synthesising or checking term, and we establish the first or second disjunct depending on~$\dir d$; otherwise, some~$t_i$, and therefore~$t$, have missing annotations, and we establish the third disjunct, applying $\ChkRule{Sub}$ to get the mode right if $\dir d = {\syn}$.
\vspace{-\topsep-\baselineskip}
\end{itemize}
\end{proof}

\Josh{Separating generalised mode preprocessing into a lemma and a theorem to simplify the case analyses}

\begin{lemma}\label{thm:Pre?-true}
If\/ $V |-_{\Sigma,\Omega} t^{\dir{d}\,\True\,e}$, then $V |-_{\Sigma,\Omega} t^\dir{d}$.
\end{lemma}

\begin{proof}
Induction on the given derivation.
The $\SynRule{Missing}$ rule cannot appear, and the other rules are mapped to their counterparts.
\end{proof}

\begin{lemma}\label{thm:Pre?-false}
If\/ $V |-_{\Sigma,\Omega} t^{\dir{d}\,\False\,e}$, then $V \not|-_{\Sigma,\Omega} t^\dir{d}$.
\end{lemma}

\begin{proof}
Induction on the generalised mode derivation, in each case analysing an arbitrary mode derivation and showing that it cannot exist.
The key case is $\SynRule{Missing}$, where we have a sub-derivation of $V |-_{\Sigma,\Omega} t^{\chk\,v\,\True}$ for some boolean~$v$.
We do not have an induction hypothesis and seem to get stuck, because $v$~is not necessarily~$\False$ --- indeed, $t$~may well be a real checking term, and this is where it is required to (but cannot) be synthesising.
But all that matters here is that $t$~is \emph{exactly} in checking mode: if we continue to analyse the sub-derivation, the outermost rule must be $\Rule{Op}$ with $\dir d = {\chk}$, implying that $t$~has to be an operation in checking mode.
Then a case analysis shows that it is impossible to have a (synthesising) mode derivation of $V |-_{\Sigma,\Omega} t^{\syn}$.
%\begin{itemize}
%\item $\SynRule{Anno}$:
%Any mode derivation must also use $\SynRule{Anno}$ as the outermost rule, and then the induction hypothesis suffices.
%\item $\ChkRule{Sub}$:
%We have a sub-derivation of $V |-_{\Sigma,\Omega} t^{\syn\,\False\,\True}$.
%Any mode derivation of $V |-_{\Sigma,\Omega} t^{\chk}$ must use $\ChkRule{Chk}$ as the outermost rule.
%(The $\Rule{Op}$ rule with $\dir d = {\chk}$ is impossible because the sub-derivation shows that $t$~is exactly a synthesising term.)
%We then have a sub-derivation of $V |-_{\Sigma,\Omega} t^{\syn}$, and can use the induction hypothesis.
%\end{itemize}
\end{proof}

\begin{corollary}[Mode Preprocessing]\label{thm:mode-preprocessing}
  It is decidable for any raw term $V |-_{\Sigma,\erase\Omega} t$ and mode~$\dir{d}$ whether $V |-_{\Sigma,\Omega} \isTerm{t}^\dir{d}$.
\end{corollary}

\begin{proof}
By \cref{thm:generalised-mode-preprocessing}, there is a derivation of $V |-_{\Sigma,\erase\Omega} t^{\dir{d}\,v\,e}$ for some booleans $v$~and~$e$, and then simply check~$v$ and apply either \cref{thm:Pre?-true} or \cref{thm:Pre?-false} to obtain $V |-_{\Sigma,\Omega} t^\dir{d}$ or $V \not|-_{\Sigma,\Omega} t^\dir{d}$.
\end{proof}

\begin{corollary}\label{thm:toPre?-true}
If\/ $V |-_{\Sigma,\Omega} t^\dir{d}$, then $V |-_{\Sigma,\Omega} t^{\dir{d}\,\True\,e}$ for some boolean~$e$.
\end{corollary}

\begin{corollary}\label{thm:toPre?-false}
If\/ $V \not|-_{\Sigma,\Omega} t^\dir{d}$, then $V |-_{\Sigma,\Omega} t^{\dir{d}\,\False\,e}$ for some boolean~$e$.
\end{corollary}

Whereas it is possible to prove \cref{thm:toPre?-true} directly by induction, converting a mode derivation to a generalised one, the negated antecedent of \cref{thm:toPre?-false} does not provide useful information.
To prove \cref{thm:toPre?-false}, basically we have to construct a generalised mode derivation from scratch for a raw term, but this is exactly generalised mode preprocessing, so we can simply reuse \cref{thm:generalised-mode-preprocessing}.
And \cref{thm:toPre?-true} can also be proved easily in the same way.

\begin{proof}[Proofs of \cref{thm:toPre?-true,thm:toPre?-false}]

By \cref{thm:generalised-mode-preprocessing}, there is a derivation of $V |-_{\Sigma,\Omega} t^{\dir{d}\,v\,e}$ for some booleans $v$~and~$e$.
If $V |-_{\Sigma,\Omega} t^\dir{d}$, then $v$ has to be $\True$ since $v = \False$ leads to a contradiction with \cref{thm:Pre?-false}.
Symmetrically, if $V \not|-_{\Sigma,\Omega} t^\dir{d}$, then $v$ has to be $\False$ so as not to contradict \cref{thm:Pre?-true}.
\end{proof}

\Josh{($1\frac{1}{2}$~pages)

• Generalised mode derivations and bidirectionalisation

• Intuitive meaning of $v$~and~$e$

• Decidability of the original mode judgements

}

\subsection{Annotatability Is Not Completeness}
\label{sec:annotatability}

\Josh{$\frac{1}{2}$~pages}

\begin{figure}
  \centering\small
  \judgbox{\isTerm{t} \sqsupseteq \isTerm{u}}{A raw term $t$ is more annotated than $u$ (for some bidirectional type system $(\Sigma, \Omega)$)}
  \begin{mathpar}
    \inferrule{\isTerm{t} \sqsupseteq \isTerm{u}}
              {(\isTerm{t} \annotate A) \sqsupseteq \isTerm{u}}\;\Rule{More}
    \and
    \inferrule{\vphantom{x : \Identifier}}
              {\isTerm{x} \sqsupseteq \isTerm{x}}
    \and
    \inferrule{\isTerm{t} \sqsupseteq \isTerm{u}}
              {(\isTerm{t} \annotate A) \sqsupseteq (\isTerm{u} \annotate A)}
    \and
    \inferrule{\isTerm{t_1} \sqsupseteq \isTerm{u_1} \quad \cdots \quad \isTerm{t_n} \sqsupseteq{u_n}}
              {\tmOpts \sqsupseteq \tmOpus}
  \end{mathpar}

  \caption{Annotation ordering between raw terms}
  \label{fig:annotation-order}
\end{figure}