### 2023-07-05 tex/BiSig.tex
[ tex ] Revised abstract & S1

\documentclass[acmsmall,screen]{acmart}

\acmJournal{PACMPL}
\acmVolume{0}
\acmNumber{0}
\acmArticle{0}
\acmMonth{7}

\ifPDFTeX
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[british]{babel}
\else
\usepackage{polyglossia}
\setdefaultlanguage{british}
\fi

\usepackage{bbold}
\newcommand{\hmmax}{0}
\newcommand{\bmmax}{0}
\usepackage{bm}
\usepackage{mathtools}
\input{mathlig}
\usepackage{tikz}
\usetikzlibrary{matrix,arrows}
\usepackage{mathpartir}
\usepackage{xspace}
\usepackage[capitalise,noabbrev]{cleveref}

\usepackage{xifthen}
\newcommand{\varcitet}[3][]{\citeauthor{#2}#3~[\ifthenelse{\isempty{#1}}{\citeyear{#2}}{\citeyear[#1]{#2}}]}

\usepackage[inline]{enumitem} % for environment enumerate*

\setlength{\marginparwidth}{1.25cm}
\usepackage[obeyFinal,color=yellow,textsize=scriptsize]{todonotes}

\usepackage{microtype}
\usepackage{subfiles}

\citestyle{acmauthoryear}
\AtEndPreamble{%
\theoremstyle{acmdefinition}
\newtheorem{remark}[theorem]{Remark}
}

\newcommand{\LT}[1]{\todo[author=LT,inline,color=orange!40]{{#1}}}
\newcommand{\Josh}[1]{\todo[author=Josh,inline,color=yellow!40]{{#1}}}

\input{type-notation} %% for type theory notation copied from HoTT Book
\input{macros}
\usepackage{mylhs2tex}

\begin{document}

\author{Liang-Ting Chen}
\email{liang.ting.chen.tw@gmail.com}
\orcid{0000-0002-3250-1331}
\author{Hsiang-Shang Ko}
\orcid{0000-0002-2439-1048}
\email{joshko@iis.sinica.edu.tw}
\affiliation{%
  \institution{Academia Sinica}
  \streetaddress{128 Academia Road}
  \city{Taipei}
  \country{Taiwan}
  \postcode{115}
}

\title{A Theory of Bidirectional Type Synthesis for Simple Types}

\begin{abstract}
%  Type checking and inference serve as the transition from parsed abstract syntax trees to well-typed programs.
  While there is much progress in designing bidirectional type systems and associated type synthesis algorithms on a case-by-case basis, the lack of a general theory makes it difficult to study the common principles and develop widely applicable techniques and practical tools such as a `type-synthesiser generator'.
  In response to this difficulty, we propose a simple yet general and constructive theory of bidirectional type synthesis.
%  which refers to the mutual use of type checking and synthesis based on bidirectional typing rules to determine the type of a program with type annotations.
  First, our theory is simple yet general as it applies to any syntax-directed bidirectional simple type system that can be specified by a signature.
%  which is analogous to a grammar in parsing.
  Second, our theory is constructive because it is not only formalised in a proof assistant based on Martin-L\"of type theory, but also formulated in a way that unifies computation and proof, resulting in a compact and practical theory---for example, a proof of the decidability of the type synthesis problem is also computationally a type synthesiser.
  Moreover, we introduce `mode preprocessing' to deal with missing type annotations, an aspect often overlooked in the literature.
  The main theorem is logically a trichotomy on parsed syntax trees and computationally a type synthesiser that may throw an exception due to missing type annotations.
  Being parametrised by a signature, our theory works as a verified type-synthesiser generator overall.
\end{abstract}

\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10011007.10011006.10011041</concept_id>
       <concept_desc>Software and its engineering~Compilers</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10011007.10011006.10011039.10011040</concept_id>
       <concept_desc>Software and its engineering~Syntax</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10011007.10011074.10011099.10011692</concept_id>
       <concept_desc>Software and its engineering~Formal software verification</concept_desc>
       <concept_significance>300</concept_significance>
       </concept>
   <concept>
       <concept_id>10003752.10003790.10011740</concept_id>
       <concept_desc>Theory of computation~Type theory</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~Compilers}
\ccsdesc[500]{Software and its engineering~Syntax}
\ccsdesc[500]{Theory of computation~Type theory}
\ccsdesc[300]{Software and its engineering~Formal software verification}

\keywords{bidirectional typing, type synthesis, type checking, type inference, language formalisation, \Agda}

\maketitle

\input{sec1-intro}
\input{sec2-outline}
\input{sec3-defs.tex}
\input{sec4-pre-synthesis.tex}
\input{sec5-synthesis.tex}
\subfile{sec6-formalisation}
\input{sec7-future.tex}

\begin{acks}
We thank Nathanael Arkor for the useful conversation and to thank the Programming Languages and Formal Methods Laboratory at Academia Sinica for the opportunity of sharing earlier ideas.

The work is supported by the Ministry of Science and Technology of Taiwan under grant MOST 109-2222-E-001-002-MY3.
\end{acks}

\bibliographystyle{ACM-Reference-Format}

\IfFileExists{reference.bib}{%
  \bibliography{reference}
}{\bibliography{ref}}

\end{document}