### 2023-07-11 tex/BiSig.tex
[ tex ] Minor changes

\documentclass[acmsmall,screen,review,anonymous]{acmart}

\acmJournal{PACMPL}
\acmVolume{0}
\acmNumber{0}
\acmArticle{0}
\acmMonth{7}

\ifPDFTeX
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[british]{babel}
\else
\usepackage{polyglossia}
\setdefaultlanguage{british}
\fi

\usepackage{subfiles}
\usepackage{bbold}
\newcommand{\hmmax}{0}
\newcommand{\bmmax}{0}
\usepackage{bm}
\usepackage{mathtools}
\input{mathlig}
%\usepackage{tikz}
%\usetikzlibrary{matrix,arrows}
\usepackage{mathpartir}

\usepackage{xspace}
\usepackage[capitalise,noabbrev]{cleveref}

% Used for typesetting Agda code
\usepackage{agda-lhs2tex}
\newcommand{\cons}[1]{\mathbf{#1}}
\newcommand{\iden}{\mathit}

\usepackage{xifthen}
\newcommand{\varcitet}[3][]{\citeauthor{#2}#3~[\ifthenelse{\isempty{#1}}{\citeyear{#2}}{\citeyear[#1]{#2}}]}

\usepackage[inline]{enumitem} % for environment enumerate*

\setlength{\marginparwidth}{1.25cm}
\usepackage[obeyFinal,color=yellow,textsize=scriptsize]{todonotes}

\usepackage{microtype}
\microtypesetup{babel}

\newcommand{\LT}[2][]{\todo[author=LT,color=orange!40,#1]{#2}}
\newcommand{\Josh}[2][]{\todo[author=Josh,color=yellow!40,#1]{#2}}

\input{type-notation} %% for type theory notation copied from HoTT Book
\input{macros}

\citestyle{acmauthoryear}

\AtEndPreamble{%
\newtheorem*{claim}{Claim}
\theoremstyle{acmdefinition}
%\newtheorem*{claimproof}{Proof of Claim}
\newenvironment{claimproof}{\par
  \renewcommand{\qedsymbol}{\footnotesize$\blacksquare$}%
  \begin{proof}[{Proof of Claim}]%
}{%
  \end{proof}%
}
\newtheorem{remark}[theorem]{Remark}
}

\author{Liang-Ting Chen}
\email{liang.ting.chen.tw@gmail.com}
\orcid{0000-0002-3250-1331}
\author{Hsiang-Shang Ko}
\orcid{0000-0002-2439-1048}
\email{joshko@iis.sinica.edu.tw}
\affiliation{%
  \institution{Academia Sinica}
  \streetaddress{128 Academia Road}
  \city{Taipei}
  \country{Taiwan}
  \postcode{115}
}

\title{A Theory of Bidirectional Type Synthesis for Simple Types}

\begin{abstract}
There has been much progress in designing bidirectional type systems and associated type synthesis algorithms, but only on a case-by-case basis.
This situation is in stark contrast to parsing, for which there have been general theories and widely applicable tools such as parser generators.
To remedy the situation, this paper develops a first theory of bidirectional type synthesis.
The whole theory works as a verified type-synthesiser generator for syntax-directed simple type systems:
From a signature that specifies a bidirectionally typed language, the theory produces a type synthesiser that decides whether an input abstract syntax tree has a typing derivation or reports that the input does not have enough type annotations.
Within the theory, we formally define soundness, completeness, and mode-correctness, which are sufficient conditions for deriving a correct type synthesiser.
We also propose a preprocessing step called `mode decoration', which helps the user to deal with missing type annotations and streamlines the theory.
The entire theory is formulated constructively, and has been formalised in the proof assistant \Agda with \AxiomK.

\end{abstract}
\begin{document}

\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10011007.10011006.10011041</concept_id>
       <concept_desc>Software and its engineering~Compilers</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10011007.10011006.10011039.10011040</concept_id>
       <concept_desc>Software and its engineering~Syntax</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10011007.10011074.10011099.10011692</concept_id>
       <concept_desc>Software and its engineering~Formal software verification</concept_desc>
       <concept_significance>300</concept_significance>
       </concept>
   <concept>
       <concept_id>10003752.10003790.10011740</concept_id>
       <concept_desc>Theory of computation~Type theory</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~Compilers}
\ccsdesc[500]{Software and its engineering~Syntax}
\ccsdesc[300]{Software and its engineering~Formal software verification}
\ccsdesc[300]{Theory of computation~Type theory}

\keywords{datatype-generic programming, language formalisation, \Agda}

\maketitle

\input{sec1-intro}
\input{sec2-outline}
\input{sec3-defs.tex}
\input{sec4-pre-synthesis.tex}
\input{sec5-synthesis.tex}
\input{example.tex}
\subfile{sec6-formalisation}
\input{sec7-future.tex}

\begin{acks}
%We thank Nathanael Arkor for the useful conversation.
We thank Kuen-Bang {Hou (Favonia)} for his comments and suggestions.
The work is supported by the \grantsponsor{NSTC}{National Science and Technology Council of Taiwan}{https://www.nstc.gov.tw/} under grant \grantnum{NSTC}{NSTC 112-2221-E-001-003-MY3}.
\end{acks}

\bibliographystyle{ACM-Reference-Format}

\IfFileExists{reference.bib}{%
  \bibliography{reference}
}{\bibliography{ref}}

\end{document}