### 2023-10-12 tex/ESOP24/BiSig.tex
[ tex ] Revise until S6

\documentclass[orivec,envcountsame,envcountsect,oribibl]{llncs}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[british]{babel}
\usepackage[numbers,sort&compress]{natbib}

\usepackage{soul}
\usepackage{subfiles}
\usepackage{xurl} % allow line breaks anywhere in a URL string
\usepackage[colorlinks,urlcolor=blue,citecolor=blue]{hyperref}

\usepackage{bbold}
\newcommand{\hmmax}{0}
\newcommand{\bmmax}{0}
\usepackage{bm}
\usepackage{mathtools}
\input{mathlig}
\usepackage{mathpartir}

\usepackage{xspace}
\usepackage[capitalise,noabbrev]{cleveref}

% Used for typesetting Agda code
\usepackage{agda-lhs2tex}
\newcommand{\cons}[1]{\mathbf{#1}}
\newcommand{\iden}{\mathit}

\usepackage{xifthen}
\newcommand{\varcitet}[3][]{\citeauthor{#2}#3~\ifthenelse{\isempty{#1}}{\cite{#2}}{\cite[#1]{#2}}}

\usepackage[inline]{enumitem} % for environment enumerate*
\newlist{inlineenum}{enumerate*}{1}
\setlist[inlineenum]{label=(\roman*)}

\setlength{\marginparwidth}{2cm}
\usepackage[obeyFinal,color=yellow,textsize=scriptsize]{todonotes}

\usepackage{microtype}
\microtypesetup{babel}

\newcommand{\LT}[2][]{\todo[author=LT,color=orange!40,#1]{#2}}
\newcommand{\Josh}[2][]{\todo[author=Josh,color=yellow!40,#1]{#2}}
\newcommand{\codefigure}{\small\setlength{\parindent}{0em}\setlength{\mathindent}{0em}\setlength{\abovedisplayskip}{0ex}\setlength{\belowdisplayskip}{0ex}}

\input{type-notation} %% for type theory notation copied from HoTT Book
\input{macros}

%\let\theorem\relax
%\let\definition\relax
%\spnewtheorem{theorem}{Theorem}[section]{\bfseries}{}
%\spnewtheorem{definition}{Definition}[section]{\bfseries}{}

%\author{Liang-Ting Chen
%\orcidID{0000-0002-3250-1331}
%\and 
%Hsiang-Shang Ko
%\orcidID{0000-0002-2439-1048}}
%
%\institute{Institute of Information Science, Academia Sinica, Taiwan}

\title{A formal treatment of bidirectional typing}

\begin{document}

\maketitle

\begin{abstract}
There has been much progress in designing bidirectional type systems and associated type synthesis algorithms, but mostly on a case-by-case basis.
This situation is in stark contrast to parsing, for which there have been general theories and widely applicable tools such as parser generators.
To remedy the situation, this paper develops a \emph{general} and \emph{formal} theory of bidirectional typing, and provides a \emph{verified} generator of \emph{proof-relevant} type synthesisers for simply typed languages: For every signature that specifies a mode-correct bidirectionally typed language, there exists a proof-relevant type synthesiser that for an input abstract syntax tree constructs a typing derivation if any, gives its refutation if not, or reports that the input does not have enough type annotations.
Soundness, completeness, and mode-correctness are studied universally for any simply typed language, which are sufficient conditions for deriving a type synthesiser.
We propose a preprocessing step called \emph{mode decoration}, which helps the user to deal with missing type annotations in an abstract syntax tree given by programmers.
The entire development is formalised in \Agda and can be further integrated with other language-formalisation frameworks.
\end{abstract}

\input{S1-intro}
\input{S2-outline}
\input{S3-defs.tex}
\input{S4-pre-synthesis}
\input{S5-synthesis}
\input{S6-example}
\subfile{S7-formalisation}
\input{S8-future}
\appendix
\input{appendix}

%\begin{acks}
%%We thank Nathanael Arkor for the useful conversation.
%We thank Kuen-Bang {Hou (Favonia)} for his comments and suggestions.
%The work is supported by the \grantsponsor{NSTC}{National Science and Technology Council of Taiwan}{https://www.nstc.gov.tw/} under grant \grantnum{NSTC}{NSTC 112-2221-E-001-003-MY3}.
%\end{acks}

\bibliographystyle{splncs04nat}

\IfFileExists{reference}{
\bibliography{reference}
}{\bibliography{ref}}

\end{document}