### 2023-10-12 tex/ESOP24/appendix.tex

%!TEX root = BiSig.tex

\section{Detailed definitions and proofs}\label{sec:defs-proofs}

\begin{definition}\label{def:syn-var}
  For an argument list $\biargs$, the set of \emph{synthesised type variables} is defined inductively by 
  \begin{align*}
    \synvar(\cdot)                   & = \emptyset  \\
    \synvar(\biargs, \chkbiarg[n+1]) & = \phantom{\fv(A_{n+1}) \cup {}} \synvar(\biargs) \\
    \synvar(\biargs, \synbiarg[n+1]) & = \fv(A_{n+1}) \cup           \synvar(\biargs).
  \end{align*}
\end{definition}